---
title: "Line_Up2_discordant_samples"
author: "dkisakye"
date: "2023-05-24"
output:
  html_document:
    toc: yes
  toc: default
---

## Objective
Create summary tables for manuscript

```{r cleanup, warning=FALSE, error=FALSE, message=FALSE}
# Clean environment
rm(list=ls(all=TRUE))

# Unload packages
if (!is.null(names(utils::sessionInfo()[["otherPkgs"]]))) pacman::p_unload("all")
```


```{r packages, warning=FALSE, error=FALSE, message=FALSE}
library(haven)
library(dplyr) 
library(tidyverse)
library(readxl)
library(writexl)
library(reshape2)
library(readr)
library(readstata13)
```

## Extract the samples from the 12-Month survey data
```{r }

Household_data_12 <- read.dta13("../../../LineUP2/Household member level data 12 month survey_FINAL.dta", convert.factors = TRUE, generate.factors = TRUE,
                         encoding = "UTF-8", fromEncoding = NULL, convert.underscore = FALSE,
                         missing.type = TRUE, convert.dates = TRUE, replace.strl = TRUE,
                         add.rownames = FALSE, nonint.factors = TRUE, select.rows = NULL,
                         select.cols = NULL, strlexport = FALSE, strlpath = ".") %>%
  filter(!is.na(uniqueid))
  
 names(Household_data_12) 
 

 # Number of participants enrolled
unique_ids<- Household_data_12 %>% 
  summarise(n_distinct_ids = n_distinct(uniqueid))

# No of ids with positive Bs
BS_pos <- Household_data_12 %>% 
  filter(BSdich == "Positive") 

pos_bs <-unique(nrow(BS_pos))

paste0("No of ids with positive Bs in LLN12:", pos_bs)



```

### Now find the discordant samples ie BS positive and RDT negative
```{r }
RDT_Neg_and_BS_pos_12 <- Household_data_12 %>%
                    select(rdtrslt, BSdich, parasitedensity, fpbarcode1)  %>%
                filter(rdtrslt == 0 & BSdich == 1)

# number of discordant
no_discordant_LLN12 <-unique(nrow(RDT_Neg_and_BS_pos_12)) 

paste0("No of discordant samples in LLN12:", no_discordant_LLN12)


# Proportion of discordant samples
proportion_discordant_12 <- nrow(RDT_Neg_and_BS_pos_12)/nrow(Household_data_12) *100
 
print(proportion_discordant_12)

```



## Extract BS positive samples from LLN24
```{r}
Household_data_24 <- read_dta("../../../LineUP2/Household member level data 24 month survey_FINAL.dta")%>%
  filter(!is.na(uniqueid))
 # Number of participants enrolled
 
Household_data_24 %>% 
  summarise(n_distinct_ids = n_distinct(uniqueid))

# No of ids with positive Bs
BS_pos_24 <- Household_data_24 %>% 
  filter(BSdich == 1) 

pos_bs_24 <-unique(nrow(BS_pos_24))

paste0("No of ids with positive Bs in LLN24:", pos_bs_24)
```

## Extract the discordant (RDT negative & Microscopy positive) samples from the 24-Month survey data
```{r}
RDT_Neg_and_BS_pos_24 <- Household_data_24 %>%
                    select(rdtrslt, BSdich, parasitedensity, fpbarcode1)  %>%
                filter(rdtrslt == 0 & BSdich == 1)


#number of discordant 
no_discordant_LLN24 <-unique(nrow(RDT_Neg_and_BS_pos_24))

paste0("No of discordant samples in LLN24:", no_discordant_LLN24)

#Proportion of discordant samples
proportion_discordant_24 <- nrow(RDT_Neg_and_BS_pos_24)/nrow(Household_data_24)

print(proportion_discordant_24) 


#save the dataset

# write_csv(RDT_Neg_and_BS_pos_24, "../../Results/Filtered_RDT_Neg_and_BS_pos_24.csv")


```
##Summarising characteristics of the LLN12 and LLN24 discordant samples
```{r}
Characteristics_discordant_12 <- Household_data_12 %>%
  filter(rdtrslt == 0 & BSdich == 1) %>% 
  select(fpbarcode1, rdtrslt, BSdich, gender, ageyrs, gender, temperature, mrccode, Region)

Characteristics_discordant_24 <- Household_data_24 %>%
  filter(rdtrslt == 0 & BSdich == 1)  %>% 
  select(fpbarcode1, rdtrslt, BSdich, gender, ageyrs, gender, temperature, mrccode, Region)

Characteristics_discordant_12_24 <- rbind(Characteristics_discordant_12, Characteristics_discordant_24)

# Import qpcr and merge in Qpcr data

var_ATS_data <-read_csv("../../data_out/166_RDT_neg_BS_pos_with_varATSqpcr_data.csv")

Characteristics_discordant_12_24 <- Characteristics_discordant_12_24 %>% 
  left_join(var_ATS_data, join_by(fpbarcode1, rdtrslt, BSdich))

# Age
Summary_age_discordant <- Characteristics_discordant_12_24 %>% 
   summarise(med_age = median(ageyrs),
             iqr_age = IQR(ageyrs))

# Breakdown age
less_or_equal_to_5 <- subset(Characteristics_discordant_12_24, ageyrs <= 5)
greater_than_5 <-subset(Characteristics_discordant_12_24, ageyrs > 5)

quantile(Characteristics_discordant_12_24$ageyrs, probs =seq(from = 0.25 ,to =0.75, by = 0.25))

# Gender of participants: 1 is male . 2 is female
Proportion_gender_12_24 <-Characteristics_discordant_12_24 %>% 
  count(gender) %>% 
  mutate(proportion = round(n/sum(n)*100))

# How many had detectable fever?
Fever_status_12_24 <- Characteristics_discordant_12_24%>%
  filter(temperature >= 37.5)

unique(nrow(Fever_status_12_24)) # 

proportion_fever <- unique(nrow(Fever_status_12_24))/nrow(Characteristics_discordant_12_24) *100

round(proportion_fever) # 

Summary_parasite_density <- Characteristics_discordant_12_24 %>% 
  mutate(Parasitemia_threshold = ifelse(Quantity <= 100, 0, 1)) %>%
  count(Parasitemia_threshold) %>% 
  mutate(Proportion = round(n/sum(n)*100))

Summary_table <- data.frame( 
N = nrow(Characteristics_discordant_12_24),
Age_less_5 = nrow(less_or_equal_to_5),
Prop_less_5 = round(nrow(less_or_equal_to_5)/nrow(Characteristics_discordant_12_24)*100),
Age_greater_5 = nrow(greater_than_5),
Prop_greater_5 = round(nrow(greater_than_5)/nrow(Characteristics_discordant_12_24)*100),
Median_age = Summary_age_discordant[1,1],
IQR_age= Summary_age_discordant[1,2],
Male = Proportion_gender_12_24[1,2],
Prop_male = Proportion_gender_12_24[1,3],
Fever = nrow(Fever_status_12_24),
Prop_fever = round(proportion_fever),
Parasitemia_less_or_equal_100 =  Summary_parasite_density[1,2],
Prop_parasitemia_less_or_equal_100 = Summary_parasite_density[1,3],
Median_parasite_density = median(Characteristics_discordant_12_24$Quantity),
IQR_parasite_density= IQR(Characteristics_discordant_12_24$Quantity),
row.names = "LN12 and 24"
  )

names(Summary_table) <- c("N","Age ≤  5", "Prop ≤  5yrs", "Age > 5","Prop >5", "Median_age", "IQR_age", "Gender (M)", "Proportion Gender", "Fever", "Proportion Fever", "Parasitemia ≤ 100", "Proportion Parasitemia ≤  100", "Median_parasite_density", "IQR_parasite_density" )


# save

  # write_csv(Summary_table, "../../data_out/Table_characteristics_LLNEUP_166_discordant_samples_04032024.csv")
```



## Obtain number of samples from both LLN12 and LLN24 survey which had both RDT and BS done
```{r}

filter_12 <- Household_data_12 %>% 
 filter(rdtdone & BSdone)

print(nrow(filter_12)) # 4836 Rdt and BS done

filter_24 <- Household_data_24 %>% 
  filter(rdtdone & BSdone )

print(nrow(filter_24))

RDT_and_bs_done_12_24 <-nrow(filter_12) +nrow(filter_24)

paste0("RDT and BS was performed on ", RDT_and_bs_done_12_24, " samples from LLN12 and LLN24 studies")
```


## Total no. of BS pos  from LLN12 and LLN24
## Total no. of RDT pos from LLN12 and LLN24
```{r}
Total_BS_pos <-pos_bs_24 + pos_bs

paste0("Total no of BS positive samples in LLN12 and LLN24 study:", Total_BS_pos)


RDT_pos_12 <- filter(Household_data_12, rdtrslt == 1 ) 


RDT_pos_24 <- filter(Household_data_24, rdtrslt == 1 )

Total_RDT_pos <- nrow(RDT_pos_12)  + nrow(RDT_pos_24)

## Totals and proportions
paste0("Total no of RDT positive samples in LLN12 and LLN24 study = ", Total_RDT_pos, " and ", round(Total_RDT_pos/RDT_and_bs_done_12_24* 100, 1), "% of the samples")


paste0("Total no of BS positive samples in LLN12 and LLN24 study = ", Total_BS_pos, " and ", round(Total_BS_pos/RDT_and_bs_done_12_24* 100, 1), "% of the samples")

## RDT pos and BS pos 

RDT_pos_BS_pos_12 <- filter( Household_data_12, rdtrslt == 1 & BSdich == 1)

RDT_pos_BS_pos_24 <- filter(Household_data_24, rdtrslt == 1 & BSdich == 1)

Total_RDT_pos_BS_pos_12_24 <- nrow(RDT_pos_BS_pos_12) + nrow(RDT_pos_BS_pos_24)


paste0("Total number of RDT Positive and Microscopy Positive samples in LLN12 and LLN24 :", Total_RDT_pos_BS_pos_12_24)
```



## More tables: Proportion of discordant samples out of the total of the BS positive samples from both LLNE12 and LLNE 24 studies
```{r}

## Proportion discordant

Prop_discordant_overall <- round((nrow(RDT_Neg_and_BS_pos_12)+ nrow(RDT_Neg_and_BS_pos_24))/Total_BS_pos * 100, 1)


paste0(" Proportion of RDT negative and BS positive samples in both LLN12 and LLN24 samples: ", Prop_discordant_overall)

```


## Random samples :
* RDT positive 
* RDT negative and BS negative
```{r}
# Import list of random samples
sample_df_RDT_pos <-read_csv("../../data_out/160_Random_sample_RDT_pos.csv") %>% 
transmute(rdtrslt, BSdich, `Study Subject` = fpbarcode1 )

sample_df_RDT_neg <- read_xlsx("../../data_out/320_Random_Samples_LN12_25102023.xlsx", sheet = "Sheet2") %>% 
transmute(rdtrslt, BSdich, `Study Subject` = fpbarcode1 )

Random_samples <- rbind(sample_df_RDT_pos, sample_df_RDT_neg)

# Merge with sampledb and get barcodes of Random list of samples

sampledb <- read_csv("../../databases/data-2024-02-05 (1).csv")

# merge sampledb with random sample list

merged_sampledb_rdt_bs <- Random_samples %>% 
  left_join(sampledb, by = join_by(`Study Subject`)) %>% 
  transmute(rdtrslt, BSdich, `Study Subject`, Barcode = as.character(Barcode))


# import list of samples on which var-ats qpcr was performed; shared by Innocent

# var_ats_performed <- read_csv("../../data_in/LN2 12 moonths samples with qPCR data..xlsx.csv")

var_ats_performed_2 <- read_xls("../../data_in/varATS_LN2_24_004_25_01_24.xls", sheet = "Results", col_names = FALSE)

var_ats_performed_2 <- var_ats_performed_2[46:142, ]

colnames(var_ats_performed_2) <- var_ats_performed_2[2, ]

var_ats_performed_2 <- var_ats_performed_2[3:97, ]

var_ats_performed_2 <- var_ats_performed_2 %>% 
  select(`Sample Name`, Quantity)
  
var_ats_performed_3 <- read_xls("../../data_in/varATS_LN2_24_005_29_01_2024.xls", sheet = "Results", col_names = FALSE)

var_ats_performed_3 <- var_ats_performed_3[46:108, ]

colnames(var_ats_performed_3) <- var_ats_performed_3[2, ]

var_ats_performed_3 <- var_ats_performed_3[3:62, ]

var_ats_performed_3 <-var_ats_performed_3 %>% 
  select(`Sample Name`, Quantity)

var_ats_bound <-rbind(var_ats_performed_2, var_ats_performed_3)

var_ats_bound <- var_ats_bound %>% 
  filter(`Sample Name` != "Neg ctrl") %>% 
  rename(Barcode = `Sample Name` ) %>% 
  mutate(Qpcr_status = 1)

# Merge Random samples with var ATS Qpcr data

rdt_bs_qpcr_data <- merged_sampledb_rdt_bs  %>% 
  left_join(var_ats_bound, by= join_by(`Barcode`))

# Identify those without qpcrdata; these should be in the LLNEUP ppcr database

no_qpcr_data <- rdt_bs_qpcr_data[is.na(rdt_bs_qpcr_data$Qpcr_status), ]
no_qpcr_data <- no_qpcr_data %>% 
  select(-Quantity)

random_samples_with_qpcr_data <- rdt_bs_qpcr_data[!is.na(rdt_bs_qpcr_data$Qpcr_status), ]

random_samples_with_qpcr_data <- random_samples_with_qpcr_data %>% 
  mutate(Quantity = ifelse(is.na(Quantity), 0, Quantity))
## Here some samples where CT values that were undetermined  are recorded as "NA" from the biorad machine. Rather than NA, making their parasite density 0.

# Import LLNEUP database

LN_UP_db <- read_csv("../../databases/Lineup2MergedQPCRResults.csv") %>% 
  transmute( Barcode = as.character(`SampleName`) , Quantity, `Study Subject` )

# Merge in LLNEUP db

qpcr_LN_UP_db <- no_qpcr_data %>% 
  left_join(LN_UP_db, by = join_by(Barcode, `Study Subject`)) %>% 
  mutate(Qpcr_status = ifelse(is.na(Quantity), 0, 1))

qpcr_data_from_LLN_db <- filter(qpcr_LN_UP_db, Qpcr_status == 1)

no_qpcr_data_final <- filter(qpcr_LN_UP_db, Qpcr_status == 0)

# save
# write_csv(no_qpcr_data_final, "../../data_out/no_qpcr_data_902204.csv")

# Lab team found the the qpcr data for the six samples above

qpcr_data_six_samples <- read_csv("../../data_in/no_qpcr_data_902204_from_lab.csv") %>% 
  transmute(rdtrslt, BSdich, `Study Subject`, Barcode, Quantity =`Quantity...10`, Qpcr_status = 1)
```

## Now bind all the df's for the random samples into one
```{r}

All_Random_samples <- rbind(random_samples_with_qpcr_data, qpcr_data_from_LLN_db, qpcr_data_six_samples)

All_Random_samples$Quantity <- round(as.numeric(All_Random_samples$Quantity), digits = 0) # round parasite density


# Ensure that we don't have any duplicates

nrow(unique(All_Random_samples)) # 320
any(duplicated(All_Random_samples$`Study Subject`)) #FALSE
any(duplicated(All_Random_samples$Barcode)) # FALSE
# Great!
```

## Summarise data for 320 random samples
```{r}

All_Random_samples_merged <- merge(All_Random_samples, Household_data_12, by.x = c("rdtrslt", "BSdich", "Study Subject" ), by.y = c("rdtrslt", "BSdich","fpbarcode1"), all.y = FALSE) 

All_Random_samples_sub <- All_Random_samples_merged %>% 
  select(rdtrslt, Region, mrccode,  BSdich, `Study Subject`, Quantity, ageyrs, temperature, gender, Barcode ) %>% 
  mutate(Quantity = round(Quantity))

summary(All_Random_samples_sub)

ggplot(All_Random_samples_sub, aes(x = Quantity)) +
  geom_histogram(bins = 30) +
  scale_x_log10()

summary(All_Random_samples_sub$Quantity)

# Let me come up with a summary for parasite density

less_than_or_equal_100_random <- subset(All_Random_samples_sub, Quantity <= 100)

n_less_100 <-nrow(less_than_or_equal_100_random) # 229

paste0( n_less_100," have parasite density below less or equal to 100, which is ", round(n_less_100/ nrow(All_Random_samples_sub)*100) ,"% of the total number of random samples" )

summary(less_than_or_equal_100_random$Quantity)

 # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 0.000   0.000   0.000   3.541   0.000  96.000 
 
ggplot(less_than_or_equal_100_random, aes( Quantity)) +
geom_histogram()

IQR(less_than_or_equal_100_random$Quantity) # 0

```



```{r}
greater_than_100 <-subset(All_Random_samples_sub, Quantity > 100)

nrow(greater_than_100) # 91

summary(greater_than_100$Quantity)

 # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    # 109     936    4967  134852   13510 4539762 

IQR(greater_than_100$Quantity) #

# Age

less_or_equal_to_5_random <- subset(All_Random_samples_sub, ageyrs <= 5)

greater_than_5_random <- subset(All_Random_samples_sub, ageyrs > 5)

# Gender
proportion_gender_random <-All_Random_samples_sub %>% 
  count(gender) %>% 
  mutate(proportion = round(n/sum(n)*100)) 
    
 # Fever 

fever_random <- subset(All_Random_samples_sub, temperature >= 37.5)

summary_table_random <- data.frame( 
N = nrow(All_Random_samples_sub),
Age_less_5 = nrow(less_or_equal_to_5_random),
Prop_less_5 = round(nrow(less_or_equal_to_5_random)/nrow(All_Random_samples_sub)*100),
Age_greater_5 = nrow(greater_than_5_random),
Prop_greater_5 = round(nrow(greater_than_5_random)/nrow(All_Random_samples_sub)*100),
Median_age = median(All_Random_samples_sub$ageyrs),
IQR_age = IQR(All_Random_samples_sub$ageyrs),
Gender = proportion_gender_random[1,2],
Prop_gender = proportion_gender_random[1,3],
Fever = nrow(fever_random),
Prop_fever = round(nrow(fever_random)/nrow(All_Random_samples_sub)* 100),
Parasitemia_less_or_equal_100 =  nrow(less_than_or_equal_100_random),
Prop_parasitemia_less_or_equal_100 = round(nrow(less_than_or_equal_100_random)/nrow(All_Random_samples_sub) * 100),
Median_parasite_density = median(All_Random_samples_sub$Quantity),
IQR_parasite_density = IQR(All_Random_samples_sub$Quantity),
row.names = "Random_samples"
)

names(summary_table_random) <- c("N","Age ≤  5", "Prop ≤  5yrs", "Age > 5","Prop >5", "Median_age", "IQR_age", "Gender (M)", "Proportion Gender", "Fever", "Proportion Fever", "Parasitemia ≤ 100", "Proportion Parasitemia < 100", "Median_parasite_density", "IQR_parasite_density " )

# save

# write_csv(summary_table_random, "../../data_out/Table_characteristics_LLNEUP_random_samples_summary_28_Feb_2024.csv")
```


Generate list of Random samples for sps PCR
```{r}
random_sample_var_ATS_neg <-All_Random_samples_sub %>% 
  filter(Quantity <= 1) 
  
# Save

write_csv(random_sample_var_ATS_neg, "../../data_out/Random_samples_varATS_neg.csv")

```

### Generate a histogram showing the causes of discordant RDT and Microscopy results
```{r}
non_pf_species <- 32
false_positive_microscopy_results <- 43
low_parasite_density_less_100 <- 31
parasite_density_more_100 <- 59
hrp2_deletions <- 1



```


```{r print_date_and_time}
Sys.time()
```

```{r print_session_info}
# sessionInfo()
devtools::session_info()
```
