---
title: "Creating Summary tables for LLINEUP2 studies"
author: "dkisakye"
date: "2023-05-24"
output:
  html_document:
    toc: yes
  toc: default
---

## Objective
Create summary tables for manuscript

```{r cleanup, warning=FALSE, error=FALSE, message=FALSE}
# Clean environment
rm(list=ls(all=TRUE))

# Unload packages
if (!is.null(names(utils::sessionInfo()[["otherPkgs"]]))) pacman::p_unload("all")
```


```{r packages, warning=FALSE, error=FALSE, message=FALSE}
library(haven)
library(dplyr) 
library(tidyverse)
library(readxl)
library(writexl)
library(reshape2)
library(readr)
library(readstata13)
library(kableExtra)
```

```{r}
Combined_lineup_data <- read.dta13("../../../LineUP2/Combined 12m and 24m CS surveys for LLINEUP2_household member level data.dta", nonint.factors = TRUE, generate.factors=TRUE)

#str(Combined_lineup_data$survey)

survey_labels <- c("12 month survey", "24 month survey") 
Combined_lineup_data$survey <- as.character(Combined_lineup_data$survey)

Combined_lineup_data$survey <- ifelse(Combined_lineup_data$survey == "1", survey_labels[1], survey_labels[2])

nrow(Combined_lineup_data)
```

### Obtain number of samples from surveys which had both RDT and BS done
```{r}
 # First, check if observations are unique
nrow(distinct(Combined_lineup_data)) # yes

RDT_BS_done <-Combined_lineup_data %>% 
  filter(rdtdone == "Yes" & BSdone == 1)

paste0("RDT and BS were performed on ", nrow(RDT_BS_done), " from the LLINEUP surveys")

```


### Extract the BS positive samples from  survey data
```{r }

# No of ids with positive BS
BS_pos <- Combined_lineup_data %>% 
  filter(BSdich == "Positive" & rdtdone == "Yes") 

pos_bs <-unique(nrow(BS_pos))

paste0("No of ids with positive BS in LLINEUP surveys:", pos_bs)

# Get a summary of how RDT results for those with positive BS

table(BS_pos$rdtrslt)

```
### Extract microscopy negative samples
```{r}
Bs_neg <-Combined_lineup_data %>% 
  filter(BSdich == "Negative" & rdtdone == "Yes") 

neg_bs <-unique(nrow(Bs_neg))

paste0("No of ids with Negative BS in LLINEUP surveys:", neg_bs)
# Get a summary of RDT results for BS negative samples
table(Bs_neg$rdtrslt)


```

### Now find the discordant samples ie BS positive and RDT negative
```{r }

Discordant<- Combined_lineup_data %>%
                filter(rdtrslt == "Negative"  & BSdich == "Positive")

# number of discordant
no_discordant <-unique(nrow(Discordant)) 

# Proportion of discordant samples out of BS positive samples
proportion_discordant_12 <- nrow(Discordant)/nrow(BS_pos) *100
 
proportion_discordant_12_rounded <-print(round(proportion_discordant_12, 2))

paste0("No of discordant samples is ", no_discordant, " and " , proportion_discordant_12_rounded,"% of all BS positive samples"  )


```

### Summarise characteristics of the discordant samples from both LLINEUP12 and LLINEUP24 surveys
```{r}
Discordant_sub <- Discordant %>% 
  select( rdtrslt, BSdich, gender, ageyrs, gender,inclusion_criteria, temperature,agecat,elevatedtemp,parasitedensity, mrccode)

# Breakdown age by category

Proportion_age_categories <-Discordant_sub %>% 
  count(agecat) %>% 
  mutate(proportion = round(n/sum(n)*100, 1))

# Summarise gender
Proportion_gender_discordant <-Discordant_sub %>% 
  count(gender) %>% 
  mutate(proportion = round(n/sum(n)*100, 1))

# How many had detectable fever?
Discordant_sub_fever <- Discordant_sub %>% 
  filter(elevatedtemp == "Yes")

# Keep observations with temp >=38.0
Discordant_sub_fever$temperature <-as.numeric(Discordant_sub_fever$temperature)

 Discordant_sub_fever<- Discordant_sub_fever%>%
  filter(temperature >= 38.0) 
 
proportion_fever <- unique(nrow(Discordant_sub_fever))/nrow(Discordant_sub) *100

proportion_fever

# Categorise parasite density by microscopy and summarise it

Summary_parasite_density_discordant <- Discordant_sub %>% 
  mutate(Parasitemia_threshold = ifelse(parasitedensity < 1000, 0,1 )) %>%
  count(Parasitemia_threshold) %>% 
  mutate(Proportion = round(n/sum(n)*100, 2))

q1<-quantile(Discordant_sub$parasitedensity, 1/4)
q1
q3 <-quantile(Discordant_sub$parasitedensity, 3/4)
q3

summary_table_discordant <-
  data.frame(
  Characteristic =c("<5", "5-15", ">=16", "Gender (M)",
  "Temperature ≥ 38.0 °C", "< 1000 parasites/μL", 
  "Median parasite density"),
  
  N = c(Proportion_age_categories[1,2], Proportion_age_categories[2,2],Proportion_age_categories[3,2], Proportion_gender_discordant[1,2],nrow(Discordant_sub_fever), Summary_parasite_density_discordant[1,2], median(Discordant_sub$parasitedensity)),

  Proportion_or_Quantiles = c(Proportion_age_categories[1,3],Proportion_age_categories[2,3], Proportion_age_categories[3,3], Proportion_gender_discordant[1,3],  round(proportion_fever, 2), Summary_parasite_density_discordant[1,3], 
  "48-2590")
)

kable(summary_table_discordant, format = "html",  
      caption = "Characteristics of discordant samples") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```


### Summarise characteristics of the concordant samples from both LLINEUP12 and LLINEUP24 surveys
```{r}
Concordant <- Combined_lineup_data %>% 
  mutate(Concordant_result = case_when(
              rdtrslt == "Positive" & BSdich == "Positive" ~ 1, TRUE ~ 0
                )) %>%
              filter(Concordant_result == 1) %>% 
                select( rdtrslt, BSdich, gender, ageyrs, gender,inclusion_criteria, temperature,agecat,elevatedtemp,parasitedensity, mrccode) 
  
# Breakdown age
Proportion_age_categories_concordant <- Concordant %>% 
  count(agecat) %>% 
  mutate(proportion = round(n/sum(n)*100, 1))

# Summarise gender
Proportion_gender_concordant <-Concordant %>% 
  count(gender) %>% 
  mutate(proportion = round(n/sum(n)*100, 1))

# How many had detectable fever?
Concordant_fever <- Concordant %>% 
  filter(elevatedtemp == "Yes")

# Keep observations with temp >=38.0
Concordant_fever$temperature <-as.numeric(Concordant_fever$temperature)

 Concordant_fever<- Concordant_fever%>%
  filter(temperature >= 38.0) 

proportion_fever_concordant <- unique(nrow(Concordant_fever))/nrow(Concordant) *100

proportion_fever

# Categorise parasite density by microscopy and summarise it

Summary_parasite_density_concordant <- Concordant %>% 
  mutate(Parasitemia_threshold = ifelse(parasitedensity < 1000, 0,1 )) %>%
  count(Parasitemia_threshold) %>% 
  mutate(Proportion = round(n/sum(n)*100, 2))

q1_concordant<-quantile(Concordant$parasitedensity, 1/4)
q1_concordant
q3_concordant <-quantile(Concordant$parasitedensity, 3/4)
q3_concordant

summary_table_concordant <-
  data.frame(
  Characteristic =c("<5", "5-15", ">=16", "Gender (M)",
  "Temperature ≥ 38.0 °C", "< 1000 parasites/μL", 
  "Median parasite density"),
  
  N = c(Proportion_age_categories_concordant[1,2], Proportion_age_categories_concordant[2,2],Proportion_age_categories_concordant[3,2], Proportion_gender_concordant[1,2],nrow(Concordant_fever), Summary_parasite_density_concordant[1,2], median(Concordant$parasitedensity)),

Proportion_or_Quantiles = c(Proportion_age_categories_concordant[1,3],Proportion_age_categories_concordant[2,3], Proportion_age_categories_concordant[3,3], Proportion_gender_concordant[1,3],  round(proportion_fever_concordant, 1), Summary_parasite_density_concordant[1,3], 
  "760-12350")
)

kable(summary_table_concordant, format = "html",  
      caption = "Characteristics of concordant samples") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


### Calculate Chi-squared test to compare observed proportions of fever between groups
```{r}
# Create a contingency table

n_discordant <- nrow(Discordant_sub)

n_fever_discordant <-nrow(Discordant_sub_fever)

n_concordant <-nrow(Concordant)

n_fever_concordant <- nrow(Concordant_fever)

discordant <- c(n_fever_discordant, n_discordant - n_fever_discordant)

concordant <- c(n_fever_concordant, n_concordant - n_fever_concordant)

fever_data <- rbind(discordant, concordant)
colnames(fever_data) <- c("Fever", "No Fever")
rownames(fever_data) <- c("Discordant", "Concordant")

# Perform chi-square test
chi_sq_test <- chisq.test(fever_data)

# Print the results
print(chi_sq_test)
```

### Perform the Mann-Whitney U/wilcoxon ranksum test to get stats on median parasite densities between concordant and discordant samples
```{r}

wilcox.test(Concordant$parasitedensity, Discordant_sub$parasitedensity, alternative = "two.sided", conf.int = TRUE)

```


```{r print_date_and_time}
Sys.time()
```

```{r print_session_info}
# sessionInfo()
devtools::session_info()
```
