---
title: "Extracting LLINE 12 and LLINE 24 sample information"
author: "dkisakye"
date: "2023-05-24"
output:
  html_document:
    toc: yes
  toc: default
---

## Objective
* Select samples that are discordant from the LLINE12 and LLINE24 studies
* Create tables summarising data from the LLNE12 and LLINE24 studies

```{r cleanup, warning=FALSE, error=FALSE, message=FALSE}
# Clean environment
rm(list=ls(all=TRUE))

# Unload packages
if (!is.null(names(utils::sessionInfo()[["otherPkgs"]]))) pacman::p_unload("all")
```


```{r packages, warning=FALSE, error=FALSE, message=FALSE}
library(haven)
library(dplyr) 
library(tidyverse)
library(readxl)
library(writexl)
library(reshape2)
library(readr)
library(readstata13)
```

## Extract the samples from the 12-Month survey data
```{r }

Household_data_12 <- read.dta13("../../../LineUP2/Household member level data 12 month survey_FINAL.dta", convert.factors = TRUE, generate.factors = TRUE,
                         encoding = "UTF-8", fromEncoding = NULL, convert.underscore = FALSE,
                         missing.type = TRUE, convert.dates = TRUE, replace.strl = TRUE,
                         add.rownames = FALSE, nonint.factors = TRUE, select.rows = NULL,
                         select.cols = NULL, strlexport = FALSE, strlpath = ".") %>%
  filter(!is.na(uniqueid))


 # Number of participants enrolled
unique_ids<- Household_data_12 %>% 
  summarise(n_distinct_ids = n_distinct(uniqueid))



```

## Now find the discordant samples ie BS positive and RDT negative
```{r }
RDT_Neg_and_BS_pos_12 <- Household_data_12 %>%
                    select(rdtrslt, BSdich, parasitedensity, fpbarcode1)  %>%
                filter(rdtrslt == "Negative" & BSdich == "Positive")

# number of discordant
no_discordant_LLN12 <-unique(nrow(RDT_Neg_and_BS_pos_12)) 

paste0("No of discordant samples in LLN12:", no_discordant_LLN12)


```



## Extract samples from LLN24
```{r}
Household_data_24 <- read.dta13("../../../LineUP2/Household member level data 24 month survey_FINAL.dta", convert.factors = TRUE, generate.factors = TRUE,
                         encoding = "UTF-8", fromEncoding = NULL, convert.underscore = FALSE,
                         missing.type = TRUE, convert.dates = TRUE, replace.strl = TRUE,
                         add.rownames = FALSE, nonint.factors = TRUE, select.rows = NULL,
                         select.cols = NULL, strlexport = FALSE, strlpath = ".") 
  
  
 # Number of participants enrolled
 
unique_ids_24 <- Household_data_24 %>% 
  summarise(n_distinct_ids = n_distinct(uniqueid))

```

## Extract the discordant samples from the 24-Month survey data
```{r}
RDT_Neg_and_BS_pos_24 <- Household_data_24 %>%
                    select(rdtrslt, BSdich, parasitedensity, fpbarcode1)  %>%
                filter(rdtrslt == "Negative" & BSdich == "Positive")


#number of discordant 
no_discordant_LLN24 <-unique(nrow(RDT_Neg_and_BS_pos_24))

paste0("No of discordant samples in LLN24:", no_discordant_LLN24)


all_discordant <- rbind(RDT_Neg_and_BS_pos_12, RDT_Neg_and_BS_pos_24)


#save the dataset

# write_csv(all_discordant, "../../data_out/Discordant_sample_list_LLINE12_LLNE24")
```

## Summarising characteristics of the LLN12 and LLN24 discordant samples
```{r}
Characteristics_discordant_12 <- Household_data_12 %>%
  filter(rdtrslt == "Negative" & BSdich == "Positive") %>% 
  select(fpbarcode1, rdtrslt, BSdich, gender, ageyrs, gender,inclusion_criteria, temperature,agecat, mrccode, Region)

Characteristics_discordant_24 <- Household_data_24 %>%
  filter(rdtrslt == "Negative" & BSdich == "Positive")  %>% 
  select(fpbarcode1, rdtrslt, BSdich, gender, ageyrs, gender, temperature, mrccode, agecat,inclusion_criteria, Region)

Characteristics_discordant_12_24 <- rbind(Characteristics_discordant_12, Characteristics_discordant_24)

Characteristics_discordant_12_24$rdtrslt <- factor(Characteristics_discordant_12_24$rdtrslt, levels = c( "Negative", "Positive"))

Characteristics_discordant_12_24$BSdich<- factor(Characteristics_discordant_12_24$BSdich, levels = c( "Negative", "Positive"))

# Convert to numeric
Characteristics_discordant_12_24$rdtrslt <- as.numeric(Characteristics_discordant_12_24$rdtrslt) - 1

Characteristics_discordant_12_24$BSdich <- as.numeric(Characteristics_discordant_12_24$BSdich) - 1

Characteristics_discordant_12_24$temperature<- as.numeric(as.character(Characteristics_discordant_12_24$temperature))


# Import qpcr and merge in Qpcr data

var_ATS_data <-read_csv("../../data_out/166_RDT_neg_BS_pos_with_varATSqpcr_data.csv")

var_ATS_data<- var_ATS_data[1:7]

Characteristics_discordant_12_24_merged <- Characteristics_discordant_12_24 %>%
  left_join(var_ATS_data, join_by(rdtrslt, BSdich, fpbarcode1))

# Age
Summary_age_discordant <- Characteristics_discordant_12_24_merged %>% 
   summarise(med_age = median(ageyrs),
             iqr_age = IQR(ageyrs))

# Quantiles
quantile(Characteristics_discordant_12_24_merged$ageyrs, probs =seq(from = 0.25 ,to =0.75, by = 0.25))

#25% 50% 75% 
 # 3   6   8 

paste0("Median_age is ", Summary_age_discordant[1,1], " and IQR is " , Summary_age_discordant[1,2])

range(Characteristics_discordant_12_24_merged$ageyrs)

ggplot(Characteristics_discordant_12_24_merged, aes(x = ageyrs)) +
  geom_histogram()

# Breakdown age

Proportion_age_categories <-Characteristics_discordant_12_24_merged %>% 
  count(agecat) %>% 
  mutate(proportion = round(n/sum(n)*100))

# Make column for new age categories < 5, 5-10, >10 years

Characteristics_discordant_12_24_merged$ageyrs <- round(Characteristics_discordant_12_24_merged$ageyrs, 0)

Characteristics_discordant_12_24_merged <- Characteristics_discordant_12_24_merged %>% 
  mutate(age_cat2 = case_when(
  ageyrs > 0 & ageyrs < 5 ~ "< 5 years",
  ageyrs >= 5 & ageyrs <= 10 ~ "5 - 10 years",
  ageyrs > 10 ~ "> 10 years"
  ))


Proportion_age_categories_2 <-Characteristics_discordant_12_24_merged %>% 
  count(age_cat2) %>% 
  mutate(proportion = round(n/sum(n)*100))

# Export summarised age categories
# write_csv(Proportion_age_categories_2, "../../data_out/Age_categories_discordant_samples.csv")

# Gender of participants: 1 is male . 2 is female
Proportion_gender_12_24 <-Characteristics_discordant_12_24_merged %>% 
  count(gender) %>% 
  mutate(proportion = round(n/sum(n)*100))

# Region
Proportion_region_12_24 <-Characteristics_discordant_12_24_merged %>% 
  count(Region) %>% 
  mutate(proportion = round(n/sum(n)*100))

# Export_ region summary
#write_csv(Proportion_region_12_24, "../../data_out/Regions_for_discordant_samples.csv")

# How many had detectable fever?
Fever_status_12_24 <- Characteristics_discordant_12_24_merged%>%
  filter(temperature >= 37.5)

unique(nrow(Fever_status_12_24)) # 

proportion_fever <- unique(nrow(Fever_status_12_24))/nrow(Characteristics_discordant_12_24_merged) *100

round(proportion_fever) # 
 
# Categorise parasite density by microscopy and summarise it


Summary_parasite_density <- Characteristics_discordant_12_24_merged %>% 
  mutate(Parasitemia_threshold = ifelse(parasitedensity < 1000, 0,1 )) %>%
  count(Parasitemia_threshold) %>% 
  mutate(Proportion = round(n/sum(n)*100))

Summary_table <- data.frame( 
N = nrow(Characteristics_discordant_12_24_merged),
Age_less_5 = Proportion_age_categories_2[2,2],
Prop_less_5 = Proportion_age_categories_2[2,3],
Age_5_to_10 = Proportion_age_categories_2[1,2],
Prop_5_to_10 = Proportion_age_categories_2[1,3],
Age_greater_10 = Proportion_age_categories_2[3,2],
Prop_greater_10 = Proportion_age_categories_2[3,3],
Median_age = Summary_age_discordant[1,1],
IQR_age= Summary_age_discordant[1,2],
Male = Proportion_gender_12_24[1,2],
Prop_male = Proportion_gender_12_24[1,3],
Fever = nrow(Fever_status_12_24),
Prop_fever = round(proportion_fever),
Parasitemia_less_1000 =  Summary_parasite_density[1,2],
Prop_parasitemia_less_1000 = Summary_parasite_density[1,3],
Median_parasite_density = median(Characteristics_discordant_12_24_merged$parasitedensity),
IQR_parasite_density= IQR(Characteristics_discordant_12_24_merged$parasitedensity),
row.names = "LN12 and 24"
  )


#names(Summary_table) <- c("N","Age ≤  5", "Prop ≤  5yrs", "Age > 5","Prop >5", "Median_age", "IQR_age", "Gender (M)", "Proportion Gender", "Fever", "Proportion Fever", "Parasitemia ≤ 100", "Proportion Parasitemia ≤  100", "Median_parasite_density", "IQR_parasite_density" )


# save


 #write_csv(Summary_table, "../../data_out/Table_characteristics_LLNEUP_166_discordant_samples_06032024.csv")
  
 quantile(Characteristics_discordant_12_24_merged$parasitedensity)
```



## Obtain number of samples from both LLN12 and LLN24 survey which had both RDT and BS done
```{r}

# convert to variables to factors
Household_data_12 <- Household_data_12 %>% 
  mutate(rdtdone = case_when(
                  rdtdone == "No" ~ 0,
                  rdtdone == "Yes" ~ 1,
                  rdtdone == "-9" ~ NA_integer_ ),
          BSdone = factor(BSdone))


filter_12 <- Household_data_12 %>% 
 filter(rdtdone == 1 & BSdone == 1)

print(nrow(filter_12)) # 3886 Rdt and BS done


Household_data_24 <- Household_data_24 %>% 
  mutate(rdtdone = case_when(
                  rdtdone == "No" ~ 0,
                  rdtdone == "Yes" ~ 1,
                  rdtdone == "-9" ~ NA_integer_),
         BSdone = factor(BSdone))

Household_data_24$BSdone <-factor(Household_data_24$BSdone)

filter_24 <- Household_data_24 %>% 
  filter(rdtdone == 1 & BSdone == 1 )

print(nrow(filter_24))

RDT_and_bs_done_12_24 <-nrow(filter_12) +nrow(filter_24)

paste0("RDT and BS was performed on ", RDT_and_bs_done_12_24, " samples from LLN12 and LLN24 studies") # 6353
```

### Total no. of BS pos  from LLN12 and LLN24
### Total no. of RDT pos from LLN12 and LLN24
```{r}

BS_pos_12 <- filter(Household_data_12, BSdich == "Positive")

BS_pos_24 <- filter(Household_data_24, BSdich == "Positive")

Total_BS_pos <- nrow(BS_pos_12) + nrow(BS_pos_24)

paste0("Total no of BS positive samples in LLN12 and LLN24 study:", Total_BS_pos)


RDT_pos_12 <- filter(Household_data_12, rdtrslt == 1 ) 


RDT_pos_24 <- filter(Household_data_24, rdtrslt == 1 )




```



## Proportion of discordant samples out of the total of the BS positive samples from both LLINE12 and LLINE 24 studies
```{r}

## Proportion discordant

Prop_discordant_overall <- round((nrow(RDT_Neg_and_BS_pos_12)+ nrow(RDT_Neg_and_BS_pos_24))/Total_BS_pos * 100, 1)


paste0(" Proportion of RDT negative and BS positive samples in both LLN12 and LLN24 samples: ", Prop_discordant_overall)

```


```{r print_date_and_time}
Sys.time()
```

```{r print_session_info}
# sessionInfo()
devtools::session_info()
```
