---
title: "Sensitivity and Specificity analysis"
author: "dkisakye/JBriggs"
date: "2024-03-25"
output:
  html_document:
    toc: yes
  toc: default
---

## Objective
Perform sensitivity and specificity analysis on samples from the LLINEUP2 study 

```{r cleanup, warning=FALSE, error=FALSE, message=FALSE}
# Clean environment
rm(list=ls(all=TRUE))

# Unload packages
if (!is.null(names(utils::sessionInfo()[["otherPkgs"]]))) pacman::p_unload("all")

```

### Load packages
```{r packages, warning=FALSE, error=FALSE, message=FALSE}
library(haven)
library(tidyverse)
library(readxl)
library(writexl)
library(reshape2)
library(readr)
library(caret)
library(pROC)
library(readstata13)
library(knitr)
library(kableExtra)

```



### Sensitivity analysis using the entire LineUP2 12 months survey samples -- JB comment i think we should combine the 12 and 24 month survey for the microscopy only analysis 
### Maybe this should go in the text as separate table - overall sensivity and specificity of RDT by microscopy in all LLN2 samples

```{r, sensitivity}

# Household_data_12 <- read.dta13("~/Documents/K23_LLINEUP2_12month_qPCR/hrp2_3_investigation/P.falciparum_HRP2_3_project-main/data_in/Household member level data 12 month survey_FINAL.dta", convert.factors = TRUE, generate.factors = TRUE,
# encoding = "UTF-8", fromEncoding = NULL, convert.underscore = FALSE,
# missing.type = TRUE, convert.dates = TRUE, replace.strl = TRUE,
# add.rownames = FALSE, nonint.factors = TRUE, select.rows = NULL,
# select.cols = NULL, strlexport = FALSE, strlpath = ".") 


 # Household_data_12 <- read_dta("~/Documents/K23_LLINEUP2_12month_qPCR/hrp2_3_investigation/P.falciparum_HRP2_3_project-main/data_in/Household member level data 12 month survey_FINAL.dta") %>% 
Household_data_12 <-read_dta("../../../LineUP2/Household member level data 12 month survey_FINAL.dta") %>% 
  select(rdtdone, rdtrslt, BSdone, BSdich) 

 # Household_data_24 <- read_dta("~/Documents/K23_LLINEUP2_12month_qPCR/hrp2_3_investigation/P.falciparum_HRP2_3_project-main/data_in/Household member level data 24 month survey_FINAL.dta") %>% 
Household_data_24 <-read_dta("../../../LineUP2/Household member level data 24 month survey_FINAL.dta") %>% 
 select(rdtdone, rdtrslt, BSdone, BSdich) 
 
 Household_data <- bind_rows(Household_data_12, Household_data_24)

# Take a look at a summary of the rdt and bs data
 table(Household_data$rdtrslt) # rdt result not available/not done for some
 
 table(Household_data$BSdich)
 
# First filter to keep only observations where rdt and bs were done -- JB comment: in 2 to 10yos only
 Household_data_trim <-Household_data %>% 
   filter(rdtdone == 1 & BSdone == 1)
 
 # Convert rdt result and blood smear microscopy result to factors

Household_data_trim$rdtrslt <- factor(Household_data_trim$rdtrslt, levels = c(1,0))

Household_data_trim$BSdich <- factor(Household_data_trim$BSdich, levels = c(1,0)) 

table(Household_data_trim$rdtrslt) 
table(Household_data_trim$BSdich) 

sensitivity_by_bs <-round(sensitivity(Household_data_trim$rdtrslt, Household_data_trim$BSdich, positive = levels(Household_data_trim$BSdich)[1]) * 100, 0) 

sensitivity_by_bs # 92

# Double check manually

true_pos_bs <- filter(Household_data_trim, rdtrslt == 1 & BSdich == 1)
 
false_neg_bs <- filter(Household_data_trim, rdtrslt == 0 & BSdich == 1 )

sensitivity_check_bs <- round(nrow(true_pos_bs)/(nrow(true_pos_bs) + nrow(false_neg_bs)) * 100, 0)

sensitivity_check_bs # Also 92

# positive predictive value of RDTs with microscopy as gold standard

pos_Pred_bs <- round(posPredValue(Household_data_trim$rdtrslt, Household_data_trim$BSdich, positive = levels(Household_data_trim$BSdich)[1])*100, 0) # 49

pos_Pred_bs

#specificity by BS, all LN2 samples
spec_by_bs <- round(specificity( Household_data_trim$rdtrslt, Household_data_trim$BSdich, negative = levels(Household_data_trim$BSdich)[-1])* 100, 0) # 57

spec_by_bs

```

## Try making a nice table of the above results
```{r, kable1} 

# Example data (replace with your data)
actual <- Household_data_trim$BSdich
predicted <- Household_data_trim$rdtrslt

# Create a confusion matrix
confusionMatrix <- table(Actual = actual, Predicted = predicted)

# Calculate Sensitivity and Specificity
TP <- confusionMatrix[1, 1]
FN <- confusionMatrix[1, 2]
TN <- confusionMatrix[2, 2]
FP <- confusionMatrix[2, 1]

sensitivity <- TP / (TP + FN)
specificity <- TN / (TN + FP)
PPV <- round(TP / (TP + FP), 2) # Positive Predictive Value
NPV <- round(TN / (TN + FN), 2) # Negative Predictive Value

# Round the values
TP <- round(TP, 2)
FN <- round(FN, 2)
TN <- round(TN, 2)
FP <- round(FP, 2)
sensitivity <- round(sensitivity, 2)
specificity <- round(specificity, 2)

# Create a data frame with all results
results <- data.frame(
  Measure = c("True Positives (TP)", "False Positives (FP)", "True Negatives (TN)", "False Negatives (FN)", 
              "Sensitivity", "Specificity", "Positive Predictive Value (PPV)", "Negative Predictive Value (NPV)"),
  Value = c(TP, FP, TN, FN, sensitivity, specificity, PPV, NPV)
)

# Create a nicely formatted table with a title
kable(results, format = "html", col.names = c("Measure", "Value"), 
      caption = "Performance Metrics Including Confusion Matrix: All LLINEUP2 samples using microscopy as gold standard") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


```
### False positivity by microscopy
```{r}
false_positivity_rate <- round(FP/(TP+FP)*100, 0)
false_positivity_rate
```



### Sensitivity analysis using the random samples with qpcr and microscopy as gold standard
### JB comment -- good to report for both qPCR and microscopy here in the random sample  

```{r}
# Import list of random samples

# Random_samples <- read_csv("~/Documents/K23_LLINEUP2_12month_qPCR/hrp2_3_investigation/P.falciparum_HRP2_3_project-main/data_in/completely_random_LN2_sample.csv")
Random_samples <- read_csv("../../data_in/completely_random_LN2_sample.csv")


# Merge Random samples with var ATS Qpcr data

# var_ATS_data <- read_csv("~/Documents/K23_LLINEUP2_12month_qPCR/hrp2_3_investigation/P.falciparum_HRP2_3_project-main/data_in/all_LN2_qpcr.csv")

var_ATS_data <- read_csv("../../data_in/all_LN2_qpcr.csv")

 All_Random_samples<- Random_samples %>% 
  left_join(var_ATS_data, by= join_by(fpbarcode1 == `Study Subject`))
 
 # Create a new variable categorical variable for qpcr result
 All_Random_samples$qpcr <- round(All_Random_samples$qpcr, 3) #rounded to 3 instead
 All_Random_samples$Qpcr_result <- ifelse(All_Random_samples$qpcr <= 1, 0, 1)
 

 # Convert variables of interest to factors
All_Random_samples$rdtrslt <- factor(All_Random_samples$rdtrslt, levels = c(1,0))
All_Random_samples$BSdich <- factor(All_Random_samples$BSdich, levels = c(1,0))
All_Random_samples$Qpcr_result <- factor(All_Random_samples$Qpcr_result, levels = c(1,0))



# Sensitivity (qPCR)
sensitivity_by_qpcr <-round(sensitivity(All_Random_samples$rdtrslt, All_Random_samples$Qpcr_result, positive = levels(All_Random_samples$Qpcr_result)[1]) * 100, 0)

sensitivity_by_qpcr # 90

# Double check manually

true_pos_qpcr <- filter(All_Random_samples, rdtrslt == 1 & Qpcr_result == 1)
 
false_neg_qpcr <- filter(All_Random_samples, rdtrslt == 0 & Qpcr_result == 1 )

sensitivity_check_qpcr <- round(nrow(true_pos_qpcr)/(nrow(true_pos_qpcr) + nrow(false_neg_qpcr)) * 100, 0)

sensitivity_check_qpcr # Also 90

#sensitivity (microscopy in random Sample)
sensitivity_by_microscopy <-round(sensitivity(All_Random_samples$rdtrslt, All_Random_samples$BSdich, positive = levels(All_Random_samples$BSdich)[1]) * 100, 0)

true_pos_bs <- filter(All_Random_samples, rdtrslt == 1 & BSdich == 1)
 
false_neg_bs <- filter(All_Random_samples, rdtrslt == 0 & BSdich == 1 )

sensitivity_check_bs <- round(nrow(true_pos_bs)/((nrow(true_pos_bs) + nrow(false_neg_bs))) * 100, 0)

sensitivity_check_bs # Also 93


# Positive predictive value of rdt with qpcr as gold standard

pos_Pred_qpcr <- round(posPredValue(All_Random_samples$rdtrslt, All_Random_samples$Qpcr_result, positive = levels(All_Random_samples$Qpcr_result)[1])*100, 0) # 64

pos_Pred_qpcr

```

### Specificity using random samples
* With Microscopy as gold standard
* With qPCR as gold standard
```{r, specificity}
# With Microscopy as gold standard

spec_by_bs <- round(specificity( All_Random_samples$rdtrslt, All_Random_samples$BSdich, negative = levels(All_Random_samples$BSdich)[-1])* 100, 0) # 56

spec_by_bs

neg_Pred_by_bs <- round(negPredValue( All_Random_samples$rdtrslt, All_Random_samples$BSdich, negative = levels(All_Random_samples$BSdich)[-1])* 100, 0) # 95 
neg_Pred_by_bs 


# Double checking manually

true_neg <- filter (All_Random_samples,  rdtrslt == 0 & BSdich == 0 )

false_pos <- filter(All_Random_samples, rdtrslt == 1 & BSdich == 0)

specificity_check_by_bs <- round(nrow(true_neg)/(nrow(true_neg) + nrow(false_pos)) * 100, 0)

specificity_check_by_bs


# With qPCR as gold standard

spec_by_qPCR <- round(specificity( All_Random_samples$rdtrslt, All_Random_samples$Qpcr_result, negative = levels(All_Random_samples$Qpcr_result)[-1])* 100, 0) # 

spec_by_qPCR # 64

# Negative predictive value by qPCR
neg_Pred_by_qPCR <- round(negPredValue( All_Random_samples$rdtrslt, All_Random_samples$Qpcr_result, negative = levels(All_Random_samples$Qpcr_result)[-1])* 100, 0) # 

neg_Pred_by_qPCR # 92
```

### Make another table 

```{r, kable2}

# Example data (replace with your data)
actual_bs <- All_Random_samples$BSdich
actual_qpcr <- All_Random_samples$Qpcr_result
predicted <- All_Random_samples$rdtrslt

# Create confusion matrices for both methods
confusionMatrix1 <- table(Actual = actual_bs, Predicted = predicted)
confusionMatrix2 <- table(Actual = actual_qpcr, Predicted = predicted)

# Calculate Sensitivity and Specificity for each method
calc_metrics <- function(confusionMatrix) {
  TP <- confusionMatrix[1, 1]
  FN <- confusionMatrix[1, 2]
  TN <- confusionMatrix[2, 2]
  FP <- confusionMatrix[2, 1]

    sensitivity <- round(TP / (TP + FN), 2)
    specificity <- round(TN / (TN + FP), 2)
    PPV <- round(TP / (TP + FP), 2) # Positive Predictive Value
    NPV <- round(TN / (TN + FN), 2) # Negative Predictive Value
    
    return(c(TP, FP, TN, FN, sensitivity, specificity, PPV, NPV))
}

# Create confusion matrices and calculate metrics for both methods
confusionMatrix1 <- table(Actual = actual_bs, Predicted = predicted)
metrics1 <- calc_metrics(confusionMatrix1)

confusionMatrix1

confusionMatrix2 <- table(Actual = actual_qpcr, Predicted = predicted)
metrics2 <- calc_metrics(confusionMatrix2)

confusionMatrix2

# Combine results into a data frame for nice formatting
results <- data.frame(
  Measure = c("True Positives (TP)", "False Positives (FP)", "True Negatives (TN)", "False Negatives (FN)", 
              "Sensitivity", "Specificity", "Positive Predictive Value (PPV)", "Negative Predictive Value (NPV)"),
  Method1 = metrics1,
  Method2 = metrics2
)

# Create a nicely formatted table with a title
kable(results, format = "html", col.names = c("Measure", "Microscopy", "qPCR"), 
      caption = "Comparison of RDT performance metrics with microscopy and qPCR as gold standard on random subset") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


```
### False positivity rate using qPCR as gold standard
```{r}

pcr_metrics <-calc_metrics(confusionMatrix2)
pcr_metrics


false_positivity_rate_qpcr <- round(pcr_metrics[2]/(pcr_metrics[1] + pcr_metrics[2]) *100, 0)

false_positivity_rate_qpcr

```

## Could be interesting to look at sensitivity of RDTs by parasite density (qPCR)

```{r}

All_Random_samples <- All_Random_samples %>% 
  mutate(qpcr_cat = ifelse(qpcr > 0, floor(log10(qpcr)), "Negative")) %>% 
  mutate(qpcr_cat1 = case_when(qpcr == 0 ~ "Negative", 
                               qpcr > 0 & qpcr <= 1 ~ "Less than or equal to 1 p/ul", 
                               qpcr > 1 & qpcr <= 10 ~ "Between 1p/uL and 10 p/uL", 
                               qpcr > 10 & qpcr <= 100 ~ "Between 10p/uL and 100 p/uL", 
                               qpcr >100 & qpcr <=1000 ~ "Between 100p/uL and 1000p/ul",
                               qpcr > 1000 & qpcr <= 10000 ~ "Between 1,000p/ul and 10,000p/ul", 
                               qpcr > 10000 ~ "Above 10,000 p/ul"))

All_Random_samples$qpcr_cat1 <- factor(All_Random_samples$qpcr_cat1, levels = c("Negative", "Less than or equal to 1 p/ul", 
                                                                                "Between 1p/uL and 10 p/uL", 
                                                                                "Between 10p/uL and 100 p/uL", 
                                                                                "Between 100p/uL and 1000p/ul", 
                                                                                "Between 1,000p/ul and 10,000p/ul", 
                                                                                "Above 10,000 p/ul"))

table(All_Random_samples$qpcr_cat1)

parasitemia_pos <- All_Random_samples %>% filter(qpcr_cat1 != "Negative")

# Replace these with your actual column names
actual_column <- 'Qpcr_result'
predicted_column <- 'rdtrslt'
parasitemia_column <- 'qpcr_cat1'

# Function to calculate sensitivity and specificity
calc_metrics <- function(data) {
  confusionMatrix <- table(data[[actual_column]], data[[predicted_column]])
  TP <- confusionMatrix[1, 1]
  FN <- confusionMatrix[1, 2]
  TN <- confusionMatrix[2, 2]
  FP <- confusionMatrix[2, 1]
  sensitivity <- TP / (TP + FN)
  specificity <- TN / (TN + FP)
  return(c(sensitivity, specificity))
}

# Calculate metrics for each parasitemia level
results <- parasitemia_pos %>%
  group_by(!!sym(parasitemia_column)) %>%
  summarise(Sensitivity = calc_metrics(cur_data())[1],
            Specificity = calc_metrics(cur_data())[2])

# View results
print(results)

table(parasitemia_pos$qpcr_cat1)

check <- parasitemia_pos %>% filter(qpcr_cat1 == "Above 10,000 p/ul" & rdtrslt == 0)


```


```{r, prevalence}
# Prevalence of malaria in study pop based on Microscopy
Prev <- (nrow(filter(All_Random_samples, BSdich == 1 )) / nrow(All_Random_samples)) 

Prev # 31%

confusionMatrix(data = All_Random_samples$rdtrslt,  reference = All_Random_samples$BSdich, positive = levels(All_Random_samples$BSdich)[1],
                prevalence = Prev)


```


```{r}

```

```{r print_date_and_time}
Sys.time()
```

```{r print_session_info}
# sessionInfo()
devtools::session_info()
```
